---
title: "Project Data Proof"
author: "Emily Han"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressMessages({
  library(haven)
  library(rio)
  library(dplyr)
  library(tidyverse)
  library(reshape2)
  library(PerformanceAnalytics)
  library(ggplot2)
  library(lessR)
  library(kableExtra)
  library(get_summary_stats)
  
})
```


Load the data for your replication project into R.  Produce the following:

1. A histogram of the dependent variable
2. A correlation matrix for the DV and IVs that the original authors included in the model you are replicating
3. A visual or tabular depiction of the missingness in the data from part (2); see p. 251-255 of the text.


# Study 2 {.tabset}
## Data {.tabset}
### Data load
```{r}
s2 <- import("data/Study2.sav")
```

### Data prep
```{r, message=FALSE, results='hide'}
s2$home= factor(s2$home, levels = c(1, 2,3,4), labels = c("home-answered","away-notanswered","home-return", "away-return"))
s2$consent= factor(s2$consent, levels = c(1, 2), labels = c("yes","no"))
s2$etnic_ascr= factor(s2$etnic_ascr, levels = c(1, 2), labels = c("slovak","roma"))
s2$etnicita= factor(s2$etnicita, levels = c(1, 2,3), labels = c("slovak","roma","other"))
s2$age <- as.numeric(s2$age)
s2 <- lessR::recode(age, new_vars="roky", old=9:72, new=1996:1933, data = s2)
s2$years <- 2019 - s2$roky
s2$gender <- factor(s2$gender, levels = c(1, 2), labels = c("Male","Female"))
s2_consent <- dplyr::filter(s2, consent=="yes")
s2_consent <- dplyr::filter(s2_consent, etnic_ascr != "NA")
s2_consent <- dplyr::filter(s2_consent, eurofondy_iv<5 & skolka_iv<5 & skolka_agree<5 & skolka_vote<5 & skolka_norms < 5 & control < 5 & suma < 5 & praca < 5 & potreba<5 & pila <5 )
```

## Descriptives {.tabset}
### Main outcomes {.tabset}
```{r}
s2_consent %>% dplyr::group_by(etnic_ascr) %>% dplyr::summarise(N = length(control), Min=min(control,na.rm= TRUE),Q1 = quantile(control,probs = .25,na.rm = TRUE),Median = median(control, na.rm = TRUE),Q3 = quantile(control,probs = .75,na.rm = TRUE),Max = max(control,na.rm = TRUE),Mean = mean(control, na.rm = TRUE),SD = sd(control, na.rm = TRUE), Skew = skewness(control, na.rm = TRUE), Kurtosis = kurtosis(control, na.rm = TRUE)) -> s2_control

knitr::kable(s2_control, caption = "Study 2 - control")%>%
  kable_styling(full_width = F)


s2_consent %>% dplyr::group_by(etnic_ascr) %>% dplyr::summarise(N = length(suma), Min=min(suma,na.rm= TRUE),Q1 = quantile(suma,probs = .25,na.rm = TRUE),Median = median(suma, na.rm = TRUE),Q3 = quantile(suma,probs = .75,na.rm = TRUE),Max = max(suma,na.rm = TRUE),Mean = mean(suma, na.rm = TRUE),SD = sd(suma, na.rm = TRUE), Skew = skewness(suma, na.rm = TRUE), Kurtosis = kurtosis(suma, na.rm = TRUE)) -> s2_equal

knitr::kable(s2_equal, caption = "Study 2 - equality")%>%
  kable_styling(full_width = F)

s2_consent %>% dplyr::group_by(etnic_ascr) %>% dplyr::summarise(N = length(praca), Min=min(praca,na.rm= TRUE),Q1 = quantile(praca,probs = .25,na.rm = TRUE),Median = median(praca, na.rm = TRUE),Q3 = quantile(praca,probs = .75,na.rm = TRUE),Max = max(praca,na.rm = TRUE),Mean = mean(praca, na.rm = TRUE),SD = sd(praca, na.rm = TRUE), Skew = skewness(praca, na.rm = TRUE), Kurtosis = kurtosis(praca, na.rm = TRUE)) -> s2_prop

knitr::kable(s2_prop, caption = "Study 2 - proportionality")%>%
  kable_styling(full_width = F)

s2_consent %>% dplyr::group_by(etnic_ascr) %>% dplyr::summarise(N = length(potreba), Min=min(potreba,na.rm= TRUE),Q1 = quantile(potreba,probs = .25,na.rm = TRUE),Median = median(potreba, na.rm = TRUE),Q3 = quantile(potreba,probs = .75,na.rm = TRUE),Max = max(potreba,na.rm = TRUE),Mean = mean(potreba, na.rm = TRUE),SD = sd(potreba, na.rm = TRUE), Skew = skewness(potreba, na.rm = TRUE), Kurtosis = kurtosis(potreba, na.rm = TRUE)) -> s2_need

knitr::kable(s2_need, caption = "Study 2 - need")%>%
  kable_styling(full_width = F)

```

```{r}

s2_consent <- mutate(s2_consent, ID = row_number())

s2_mains <- melt(s2_consent,
        # ID variables - all the variables to keep but not split apart on
    id.vars=c("etnic_ascr", "ID","eurofondy_iv","skolka_iv" ),
        # The source columns
    measure.vars=c("control", "suma", "praca","potreba","pila"),
        # Name of the destination column that will identify the original
        # column that the measurement came from
    variable.name="condition",
    value.name="measurement"
)

s2_mains <- filter(s2_mains, condition!="pila") 

s2_dvs <- s2_mains %>%
  group_by(etnic_ascr,condition)%>%
  dplyr::select(-ID:-skolka_iv) %>%
  get_summary_stats(type = c("mean_sd"))%>%
  dplyr::select(-variable)%>%
  dplyr::mutate(condition = dplyr::recode(condition,
    "control" = "Control",
    "suma" = "Equality",
    "praca" = "Reciprocity",
    "potreba" = "Need"))%>%
  dplyr::mutate(sd = round(sd,2))
      

dvstt <- s2_mains %>% 
#  pivot_longer(contains("att_"))%>% 
    group_by(etnic_ascr)%>% 
    t_test(measurement~condition, detailed = T)      

```


### Sample characteristics
```{r}
s2_consent_gender <- s2_consent %>% group_by(gender) %>% 
dplyr::summarise(n = n()) %>%
  mutate(freq = n / sum(n))
s2_consent_gender$freq <- round(s2_consent_gender$freq, 3)

s2_consent_gender%>%
kable(caption = "Gender")%>%
  kable_styling(full_width = F)

s2_consent_age <- s2_consent %>% dplyr::summarise(Min=min(years,na.rm= TRUE),Q1 = quantile(years,probs = .25,na.rm = TRUE),Median = median(years, na.rm = TRUE),Q3 = quantile(years,probs = .75,na.rm = TRUE),Max = max(years,na.rm = TRUE),Mean = round(mean(years, na.rm = TRUE),3),SD = round(sd(years, na.rm = TRUE),2),n = n(),Missing = sum(is.na(years))) 

s2_consent_age %>%
  kable(caption = "Age") %>%
  kable_styling(full_width = F)

s2_consent_etnic <- s2_consent %>% group_by(etnic_ascr) %>% 
dplyr::summarise(n = n()) %>%
  mutate(freq = n / sum(n))
s2_consent_etnic$freq <- round(s2_consent_etnic$freq, 3)

s2_consent_etnic%>%
kable(caption = "Ascribed ethnicity")%>%
  kable_styling(full_width = F)

s2_consent_etnic_self <- s2_consent %>% group_by(etnicita) %>% 
dplyr::summarise(n = n()) %>%
  mutate(freq = n / sum(n))
s2_consent_etnic_self$freq <- round(s2_consent_etnic_self$freq, 3)

s2_consent_etnic_self%>%
kable(caption = "Self-reported ethnicity")%>%
  kable_styling(full_width = F)
```



```{r}

```

## Plots {.tabset}
### Main outcomes
```{r}


#fig <- image_graph(width = 1400, height = 650, res = 192)
    
ggplot(data = s2_mains, aes(y = measurement, x = condition, fill = etnic_ascr)) + geom_flat_violin(data = s2_mains, position = position_nudge(x = .2, y = 0), alpha = .8) + geom_point(data = s2_mains, aes(y = measurement, color = etnic_ascr), position = position_jitter(width = .15,height = .15), size = .5, alpha = 1.2) +
    scale_color_grey() +
    scale_fill_grey(labels=c("Slovak","Roma")) +
    theme_bw() + raincloud_theme + labs(y = "Agreement with building social housing") + scale_x_discrete(expand = c(0.15,0.2),labels=c("Control","Equality","Reciprocity","Need"))+ labs(fill='Ascribed \nethnicity') + guides(color = FALSE) 

#dev.off()  
#out <- image_crop(fig, "1380x750+0")
#print(out)
```
